### 2016-06-30 stitched.tex
ENH:  More revisions.

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
    \setmainfont{Georgia}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows. Herein, we develop a
concatenated random forest framework with image features for segmenting
WMHs in a TBI cohort. The framework is built upon the Advanced
Normalization Tools (ANTs) and ANTsR toolkits. MR (3D FLAIR, T2-, and
T1-weighted) images from 24 service members and veterans scanned in the
Chronic Effects of Neurotrauma Consortium's (CENC) observational study
were acquired. Manual annotations were employed for both training and
evaluation using a leave-one-out strategy. Lesion load and overlap
evaluative comparisons are complimented by feature rankings which
showcase the utility of the concatenated approach. Our findings suggest
supervised learning methods may be applied to quantify WMHs on routine
brain imaging. Paired with correlative outcome data, supervised learning
methods may allow for identification of imaging features predictive of
diagnosis and prognosis in individual TBI patients.

@@

\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient}\textsuperscript{2,6--8}\textcolor{blue}{. Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury}\textsuperscript{6}\textcolor{blue}{.
Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
classify ``outcome'' in TBI patients where ``outcome'' refers to the
spectrum of possible prognoses from death to disability to recovery.}\textsuperscript{4}\textcolor{blue}{. Additionally,
in patients who are comatose following severe TBI
the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome}\textsuperscript{7}.

@@

@@

\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
subjects.  J. R. S. is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.  All
multi-modal MR dicom image slices were converted to the nifti file format}.\footnote{\url{http://nifti.nimh.nih.gov/nifti-1}}
\textcolor{blue}{Given the
utility of the FLAIR sequence in detecting white matter lesions in TBI}\textsuperscript{3}\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software}\textsuperscript{22}
\textcolor{blue}{.  The normalized MRI volumes were then provided to
J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{23}
\textcolor{blue}{
which has multi-image overlay capabilities for visualizing all modalities in all three
canonical views.}

@@

\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
include offline generation of symmetric multimodal templates, the creation of
feature images from the training data which are then employed for statistical
prediction using a concatenated random forest framework.  This framework involves
the use of two random forest models where the outcome (i.e., the tissue
probability estimates) of the first RF model application,
which we denote as ``Stage1 '', is used as input (along with the original set of
feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
Once these offline steps
are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}

@@

@@

@@

As mentioned previously, input for each subject comprises FLAIR, T1-,
and T2-weighted acquisitions. The T1 and T2 images are rigidly
registered to the FLAIR image using the open-source Advanced
Normalization Tools (ANTs)\textsuperscript{22}. The aligned images are
then preprocessed using the denoising algorithm of\textsuperscript{28}
followed by N4 bias correction\textsuperscript{29} which are then
normalized to the intensity range \([0,1]\). Although we could have used
an alternative intensity standardization algorithm
(e.g.,\textsuperscript{30}), we found that a simple linear rescaling
produced better results similar to previous work\textsuperscript{17}.

@@

\textcolor{blue}{Several measures have been employed in the literature for evaluating
automated white matter lesion segmentation involving such quantities as lesion load,
voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}\textsuperscript{37}\textcolor{blue}{.  For this work, due to the relatively small-size distribution}\footnote{\textcolor{blue}{For comparison, the training data set of
  the MS Lesion Segmentation challenge
  associated with the international MICCAI 2008 conference has a mean lesion load of
  204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
  in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}}
\textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.  The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.  It should be noted that the number of true negatives is not readily
incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments. The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}

@@

\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}

@@

\textcolor{blue}{In Figure 8 we provide the segmentation evaluations derived from the leave-one-out evaluation
of the previously described TBI data.  These performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
size ranges over which these measures are computed to illustrate the variation
in performance with lesion size.  Smaller lesions ($< 12$ voxels) are more difficult
to identify which is why the sensitivity is widely varied over the cohort vs. the largest
set of voxels ($> 28$ voxels).   The first three measures are based on the identification of
entire lesions.  The relative volume difference provides a direct assessment of how accurate
the volumetric estimate is when comparing the manually identified lesions versus the
automatically predicted lesions.}

@@

The resulting rankings for both Stages are given in Figures 5 and 6
where the values for the separate stages are averaged over the entire
corresponding model set. In addition, we track the variance for each
feature over all models to illustrate the stability of the chosen
features during the evaluation. This latter information is illustrated
as horizontal errors bars providing the \(95^{th}\) percentile Note that
the reader can cross reference Table 1 for identifying corresponding
feature types and names.

Additionally, it is interesting to note some of the other top performing
features for Stage 1. The contralateral difference FLAIR image is highly
discriminative over the set of evaluation random forest models (see
Figure 7). This accords with the known clinical relevance of FLAIR
images for identifying white matter hyperintensities and the fact that
such pathology does not typically manifest symmetrically in both
hemispheres. Interestingly, the posterior maps for the deep gray matter
are extremely important for accurate white matter hyperintensity
segmentation. Perhaps the spatial specification of deep gray matter aids
in the removal of false positives. Inspection of the bottom of the plots
demonstrates the lack of discriminating features associated with the T1
image which is also well-known in the clinical literature.

As described earlier, for Stage 2, we used the output random forest
voting maps from Stage 1 as both features themselves and as priors for
input to a Bayesian-based segmentation with an additional MRF spatial
prior. In Figure 6, the voting maps are labeled as
``\emph{RFStage1VotingMaps}'' where the final numeral is associated with
the brain parenchymal labeling given previously. Similarly, the
additional RF prior segmentation feature probability maps are labeled as
``\emph{RFBrainSegmentationPosteriors}''. The Stage 2 feature importance
plot follows similar trends as that for Stage 1 with the T1 images not
contributing much to the identification of white matter hyperintensity
voxels. The initial voting maps from Stage 1 are extremely important
with the top 3 being the estimated locations of the 1) gray matter, 2)
white matter, and 3) white matter hyperintensities. Since these tissue
type can be conflated based on intensity alone it is intuitive that such
features would be important.

@@

@@

\section{References}\label{references}

\section{Figure Captions}\label{figure-captions}

\textbf{Figure 1:}
\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.}

\textbf{Figure 2:} Canonical views of the mutlivariate, bilaterally
symmetric template constructed from the MMRR data
set\textsuperscript{25} (only shown are the FLAIR, T1, and T2
modalities--- the components relevant for this work). Template
construction is detailed in\textsuperscript{17}. These images are
important for asymmetry-based features.

\textbf{Figure 3:} Representation of Stage 1 feature images for subject
01C1019. The FLAIR, T1-, and T2-weighted images are rigidly
pre-aligned\textsuperscript{22} to the space of the T1 image. The three
modality images are then preprocessed (N4 bias
correction\textsuperscript{29} and adaptive
denoising\textsuperscript{28}) followed by application of standard ANTs
brain extraction and \(n\)-tissue segmentation protocols using the MMRR
symmetric template and corresponding priors\textsuperscript{24} applied
to the T1 image. The feature images are then generated for voxelwise
input to the RF model which results in the voting maps illustrated on
the right. This gives a probabilistic classification of tissue type. Not
shown are the probability and voting images for the brain stem and
cerebellum.

\textbf{Figure 4:} Sample FLAIR acquisition image slices showing both
manual and random forest segmentations for both stages obtained during
the leave-one-out evaluation. Manual segmentations were performed by one
of the authors and provided the ground truth WMH labels for training the
random forest models.

\textbf{Figure 5:} Average \emph{MeanDecreaseAccuracy} plots generated
from the creation of all 24 random forest models for Stage 1 during the
leave-one-out evaluation. These plots are useful in providing a
quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance. The horizontal
error bars provide the \(95^{th}\) percentile and illustrate the
stability of the feature importance across the leave-one-out models. At
this initial stage only 31 feature images are used.

\textbf{Figure 6:} Average \emph{MeanDecreaseAccuracy} plots generated
from the creation of all 24 random forest models for Stage 2 during the
leave-one-out evaluation. These plots are useful in providing a
quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance. The horizontal
error bars provide the \(95^{th}\) percentile and illustrate the
stability of the feature importance across the leave-one-out models. We
augment the 31 feature images from the first stage by adding an
additional seven voting maps and 7 segmentation posteriors from
application of the Bayesian-based segmentation for a total of 45 images
for the second stage.

\textbf{Figure 7:} (a) FLAIR image slice illustrating WMHs which have
been manually delineated. The region around the WMHs is enlarged (b) in
the original FLAIR and the (c) contralateral FLAIR difference image.

\textbf{Figure 8:} Voxelwise comparison with manual delineation of white
matter hyperintensities. On the left are the calculated Dice values over
all white matter hyperintensities. Note the improvement in the Dice
metric from the employment of the Stage 2 component of the processing
pipeline. (Right) Similar results can be seen by comparing the total
lesion load volume between manual and automated detection strategies.
Although some outliers are found after the Stage 2 processing in a
couple subjects, the number of outliers caused by false positives is
decreased significantly with the second stage processing.

@@

\hypertarget{ref-Manjon:2010aa}{}
28. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
non-local means denoising of mR images with spatially varying noise
levels. J Magn Reson Imaging. 2010;31(1):192--203.

\hypertarget{ref-Tustison:2010ac}{}
29. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
2010;29(6):1310--20.

@@

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
    \setmainfont{Georgia}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows. Herein, we develop a
concatenated random forest framework with image features for segmenting
WMHs in a TBI cohort. The framework is built upon the Advanced
Normalization Tools (ANTs) and ANTsR toolkits. MR (3D FLAIR, T2-, and
T1-weighted) images from 24 service members and veterans scanned in the
Chronic Effects of Neurotrauma Consortium's (CENC) observational study
were acquired. Manual annotations were employed for both training and
evaluation using a leave-one-out strategy.
\textcolor{blue}{Performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.
Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
relative volume difference} = 43 \(\pm\) 26\%. In addition, three lesion
size ranges are selected to illustrate the variation in performance with
lesion size.\} Paired with correlative outcome data, supervised learning
methods may allow for identification of imaging features predictive of
diagnosis and prognosis in individual TBI patients.

@@

\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient}\textsuperscript{2,6--8}\textcolor{blue}{. Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury}\textsuperscript{6}\textcolor{blue}{.
Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
classify ``outcome'' in TBI patients where ``outcome'' refers to the
spectrum of possible prognoses from death to disability to recovery}\textsuperscript{4}\textcolor{blue}{. Additionally,
in patients who are comatose following severe TBI
the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome}\textsuperscript{7}.

@@

\input{trainingData.tex}

@@

\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
subjects.  J. R. S. is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.  All
multi-modal MR dicom image slices were converted to the nifti file format}.\footnote{\url{http://nifti.nimh.nih.gov/nifti-1}}
\textcolor{blue}{All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software}\textsuperscript{22}
\textcolor{blue}{.  The normalized MRI volumes were then provided to
J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{23}
\textcolor{blue}{
which has multi-image overlay capabilities for visualizing all modalities in all three
canonical views.}

@@

\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
include offline generation of symmetric multimodal templates, the creation of
feature images from the training data which are then employed for statistical
prediction using a concatenated random forest framework.  This framework involves
the use of two random forest models where the outcome (i.e., the tissue
probability estimates) of the first RF model application,
which we denote as ``Stage1'', is used as input (along with the original set of
feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
Once these offline steps
are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}

\begin{figure}[htbp]
\centering
\includegraphics{Figures/wmhPipeline.png}
\caption{\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.}}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{Figures/MMRR.png}
\caption{Canonical views of the mutlivariate, bilaterally symmetric
template constructed from the MMRR data set\textsuperscript{25} (only
shown are the FLAIR, T1, and T2 modalities--- the components relevant
for this work). Template construction is detailed
in\textsuperscript{17}. These images are important for specific
asymmetry-based features.}
\end{figure}

@@

\input{featureImagesTable.tex}

@@

\begin{figure}[htbp]
\centering
\includegraphics{Figures/featureImages.png}
\caption{Representation of Stage 1 feature images for subject 01C1019.
The FLAIR, T1-, and T2-weighted images are rigidly
pre-aligned\textsuperscript{22} to the space of the FLAIR image. The
three modality images are then preprocessed (N4 bias
correction\textsuperscript{28} and adaptive
denoising\textsuperscript{29}) followed by application of standard ANTs
brain extraction and \(n\)-tissue segmentation protocols using the MMRR
symmetric template and corresponding priors\textsuperscript{24} applied
to the T1 image. The feature images are then generated for voxelwise
input to the RF model which results in the voting maps illustrated on
the right. This gives a probabilistic classification of tissue type. Not
shown are the probability and voting images for the brain stem and
cerebellum.}
\end{figure}

As mentioned previously, input for each subject comprises FLAIR, T1-,
and T2-weighted acquisitions. The T1 and T2 images are rigidly
registered to the FLAIR image using the open-source Advanced
Normalization Tools (ANTs)\textsuperscript{22}. The aligned images are
then preprocessed using the denoising algorithm of\textsuperscript{29}
followed by N4 bias correction\textsuperscript{28} which are then
normalized to the intensity range \([0,1]\). Although we could have used
an alternative intensity standardization algorithm
(e.g.,\textsuperscript{30}), we found that a simple linear rescaling
produced better results similar to previous work\textsuperscript{17}.

\begin{figure}[htbp]
\centering
\includegraphics{Figures/sampleResults.png}
\caption{Sample FLAIR acquisition image slices showing both manual and
random forest segmentations for both stages obtained during the
leave-one-out evaluation. Manual segmentations were performed by one of
the authors and provided the ground truth WMH labels for training the
random forest models.}
\end{figure}

@@

\textcolor{blue}{Several measures have been employed in the literature for evaluating
automated white matter lesion segmentation involving such quantities as lesion load,
voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}\textsuperscript{37}\textcolor{blue}{.  For this work, due to the relatively small-size distribution}
\textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.  The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.  It should be noted that the number of true negatives is not readily
incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments. The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}

@@

\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}

\begin{itemize}
\item
  \textcolor{blue}{Quantile 1: $[1-12)$,}
\item
  \textcolor{blue}{Quantile 2: $[12-28)$, and}
\item
  \textcolor{blue}{Quantile 3: $[28-551]$}.
\end{itemize}

\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}

@@

\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
of the previously described TBI data over the three lesion volume ranges.
These performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
size ranges over which these measures are computed are meant to illustrate the variation
in performance with lesion size.}

\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
to identify which is why the sensitivity for this range is more varied compared with the largest
set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
the volumetric estimate when comparing the manually identified lesions versus the
automatically predicted lesions.}

\textcolor{blue}{
We averaged these measures over all lesions and all subjects which resulted in the
following values:}

\begin{itemize}
\item
  \textcolor{blue}{Stage 1}

  \begin{itemize}
  \item
    \textcolor{blue}{sensitivity} = 0.70 \(\pm\) 0.34
  \item
    \textcolor{blue}{PPV} = 0.42 \(\pm\) 0.36
  \item
    \textcolor{blue}{$F_1$} = 0.47 \(\pm\) 0.36
  \item
    \textcolor{blue}{relative volume difference} = 43 \(\pm\) 38\%
  \end{itemize}
\item
  \textcolor{blue}{Stage 1}

  \begin{itemize}
  \item
    \textcolor{blue}{sensitivity} = 0.68 \(\pm\) 0.38
  \item
    \textcolor{blue}{PPV} = 0.51 \(\pm\) 0.40
  \item
    \textcolor{blue}{$F_1$} = 0.52 \(\pm\) 0.36
  \item
    \textcolor{blue}{relative volume difference} = 43 \(\pm\) 26\%
  \end{itemize}
\end{itemize}

\begin{figure}[htbp]
\centering
\includegraphics{Figures/evaluationResults.png}
\caption{\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
volume difference.   These quantitative assessments are given for three quantile ranges
spanning the range of the manually-derived lesion volumes.    Overall improvement in all
three whole lesion-based measuers is seen as the
second Stage RF model is applied for all three quantile ranges.
The relative volume difference corresponding to the Stage 2 results tend to predict a
decreased predicted volume over the Stage 1 results.}}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{Figures/averageLeaveOneOutStage1.png}
\caption{Average \emph{MeanDecreaseAccuracy} plots generated from the
creation of all 24 random forest models for Stage 1 during the
leave-one-out evaluation. These plots are useful in providing a
quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance. The horizontal
error bars provide the \(95^{th}\) percentile and illustrate the
stability of the feature importance across the leave-one-out models. At
this initial stage only 31 features images are used.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{Figures/averageLeaveOneOutStage2.png}
\caption{Average \emph{MeanDecreaseAccuracy} plots generated from the
creation of all 24 random forest models for Stage 2 during the
leave-one-out evaluation. These plots are useful in providing a
quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance. The horizontal
error bars provide the \(95^{th}\) percentile and illustrate the
stability of the feature importance across the leave-one-out models. We
augment the 31 feature images from the first stage by adding an
additional 7 voting maps and 7 segmentation posteriors from application
of the Bayesian-based segmentation for a total of 45 images for the
second stage.}
\end{figure}

The resulting rankings for both Stages are given in Figures 6 and 7
where the values for the separate stages are averaged over the entire
corresponding model set. In addition, we track the variance for each
feature over all models to illustrate the stability of the chosen
features during the evaluation. This latter information is illustrated
as horizontal errors bars providing the \(95^{th}\) percentile Note that
the reader can cross reference Table 1 for identifying corresponding
feature types and names.

@@

\textcolor{blue}{
In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies}\textsuperscript{39,40}
\textcolor{blue}{exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
were found during a literature search and evaluated which concern automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}\textsuperscript{41}.

Regarding the feature rankings, it is interesting to note some of the
other top performing features for Stage 1. The contralateral difference
FLAIR image is highly discriminative over the set of evaluation random
forest models (see Figure 8). This accords with the known clinical
relevance of FLAIR images for identifying white matter hyperintensities
and the fact that such pathology does not typically manifest
symmetrically in both hemispheres. Interestingly, the posterior maps for
the deep gray matter are extremely important for accurate white matter
hyperintensity segmentation. Perhaps the spatial specification of deep
gray matter aids in the removal of false positives. Inspection of the
bottom of the plots demonstrates the lack of discriminating features
associated with the T1 image which is also well-known in the clinical
literature.

\begin{figure}[htbp]
\centering
\includegraphics{Figures/FLAIRcontralaleteralWithLesionsAlternate.png}
\caption{(a) FLAIR image slice illustrating WMHs which have been
manually delineated. The region around the WMHs is enlarged (b) in the
original FLAIR and the (c) contralateral FLAIR difference image.}
\end{figure}

As described earlier, for Stage 2, we used the output random forest
voting maps from Stage 1 as both features themselves and as priors for
input to a Bayesian-based segmentation with an additional MRF spatial
prior. In Figure 7, the voting maps are labeled as
``\emph{RFStage1VotingMaps}'' where the final numeral is associated with
the brain parenchymal labeling given previously. Similarly, the
additional RF prior segmentation feature probability maps are labeled as
``\emph{RFBrainSegmentationPosteriors}''. The Stage 2 feature importance
plot follows similar trends as that for Stage 1 with the T1 images not
contributing much to the identification of white matter hyperintensity
voxels. The initial voting maps from Stage 1 are extremely important
with the top 3 being the estimated locations of the 1) gray matter, 2)
white matter, and 3) white matter hyperintensities. Since these tissue
type can be conflated based on intensity alone it is intuitive that such
features would be important.

\section{Conclusions}\label{conclusions}

@@

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

@@

\hypertarget{ref-Tustison:2010ac}{}
28. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
2010;29(6):1310--20.

\hypertarget{ref-Manjon:2010aa}{}
29. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
non-local means denoising of mR images with spatially varying noise
levels. J Magn Reson Imaging. 2010;31(1):192--203.

@@

\hypertarget{ref-vanginneken:2007aa}{}
41. Ginneken BV, Heimann T, Styner M. 3D segmentation in the clinic: A
grand challenge. In: 3D segmentation in the clinic: A grand challenge.
2007. pp. 7--15.
